---
title: "simpleRegress"
author: "Kimberly Orantes"
date: "Fall 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
bears$MONTH<-as.factor(bears$MONTH)
bears$SEX<-as.factor(bears$SEX)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

## Model LENGTH ~ WEIGHT  this is an example  

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

The residual standard of error for NECK variable is 5.376 and for the new comparison with the variable WEIGHT is 5.432. The adjusted R-squared for NECK is 0.7476 and the adjusted R-squared for WEIGHT is 0.7423. This shows that NECK is better than weight to predict LENGTH.

## Model LENGTH ~ CHEST  this is an example  

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted R-squared for CHEST is 0.7869 which is better than the NECK variable adjusted R-squared which is 0.7476. This shows that the CHEST is better than NECK to predict LENGTH.