---
title: "simpleRegress"
author: "Kimberly Orantes"
date: "Fall 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
bears$MONTH<-as.factor(bears$MONTH)
bears$SEX<-as.factor(bears$SEX)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

## Model LENGTH ~ WEIGHT  this is an example  

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

The residual standard of error for NECK variable is 5.376 and for the new comparison with the variable WEIGHT is 5.432. The adjusted R-squared for NECK is 0.7476 and the adjusted R-squared for WEIGHT is 0.7423. This shows that NECK is better than weight to predict LENGTH.

## Model LENGTH ~ CHEST  this is an example  

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted R-squared for CHEST is 0.7869 which is better than the NECK adjusted R-squared which is 0.7476. This shows that the CHEST is better than NECK to predict LENGTH.

## Model LENGTH ~ AGE  this is an example  

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```
The variable AGE adjusted R-squared is 0.5073 which is worse than NECK adjusted R-squared which is 0.7476. This shows AGE is not better to predict LENGTH.

## Model LENGTH ~ HEADWTH  this is an example  

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted R-squared of HEADWTH (head width) is 0.5321 which is worse than the NECK adjusted R-squared which is 0.7476, therefore is isn't better to predict LENGTH.

## Model LENGTH ~ HEADLEN  this is an example  

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted R-squared for HEADLEN (head length) is 0.8434 which is better than NECK adjusted R-squared which is 0.7476. It is also better than the adjusted R-squared of CHEST which is 0.7869. This shows that the HEADLEN variable is the best variable to predict the LENGTH of bears. The order of variables from greatest to least to predict bear LENGTH are: HEADLEN, CHEST, NECK, WEIGHT, HEADWTH, and AGE. 